---
title: Neural Network Image Classification
author: Brandon P. Pipher
date: '2020-04-14'
slug: []
categories:
  - Machine Learning
  - Image Classification
tags:
  - keras
  - R
  - tensorflow
subtitle: 'Utilizing keras in R on MNIST-style data sets'
summary: 'Utilizing keras in R on MNIST-style data sets'
authors: []
lastmod: '2020-04-14T14:09:00-04:00'
featured: no
image:
  caption: 'Entacmaea Quadricolor'
  focal_point: ''
  preview_only: no
projects: []
draft: true
output:
  blogdown::html_page:
    toc: true
---

```{r, echo = FALSE}
# https://github.com/rstudio/rmarkdown/pull/1602
knitr::opts_chunk$set(class.source='fold-show', message = FALSE, warning = FALSE, echo = TRUE)
```

# Keras and beyond
Keras is a high-level neural networks API developed with a focus on enabling fast experimentation.  Originally written in Python, the library `keras` is part of a series of R interfaces to TensorFlow.  

TensorFlow is an open-source machine learning framework from Google which includes the Neural Network library of Keras.

The goal in this post is to do some low level examples on MNIST-like data sets and act as an introduction in applying neural network classifications techniques.  For the sake of brevity this post is short on the details with the intention of being revisited.

# MNIST
The [MNIST](https://en.wikipedia.org/wiki/MNIST_database) database is composed of handwritten digits and is a common "Hello, World!" data set for neural networks.


```{r, message = FALSE, warning = FALSE}
library(keras)
library(tidyverse)
library(ggplot2)
# Used for side-by-side ggplots (similar to par(mfrow = c( , )))
library(gridExtra)
```

```{r, cache = TRUE, warning = FALSE}
mnist <- dataset_mnist()
```

Taking a look at an example of the images:

```{r, error = FALSE, message = FALSE, cache = TRUE}
# Plotting helper; mnist is global var
mnistplot <- function(picked){
  mnist$train$x[picked,,] %>% 
  # Wrangle
  as_tibble() %>% 
  rowid_to_column("Y") %>% 
  pivot_longer(-Y,names_to = "X", values_to = "PIX") %>% 
  mutate(X = as.numeric(gsub("V","",X))) %>%
  mutate(Y = abs(Y - max(Y)+1)) %>%
  # Graph
  ggplot(aes(X,Y, fill = PIX)) + 
  geom_tile() + 
  scale_fill_gradient(low="white", high="black", guide = FALSE) +
  ggtitle(paste("Handwritten number:",mnist$train$y[picked])) +
  theme_void()
}
# Grid graph
pickcount <- 9
mnistex <- lapply(1:pickcount, mnistplot)
mnistexplot <- arrangeGrob(grobs = mnistex,
                           ncol = 3,
                           nrow = 3)
plot(mnistexplot)
```

```{r}
mnist.train = mnist$train
mnist.test  = mnist$test
# Normalization; Speeds training time
mnist.train$x <- mnist.train$x/255
mnist.test$x  <- mnist.test$x/255
```

```{r}
mnist.model <- keras_model_sequential() %>%
  layer_flatten(input_shape = c(28,28)) %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(10, activation = "softmax")
```

```{r}
summary(mnist.model)
```
```{r}
mnist.model %>% 
  compile(
    loss = "sparse_categorical_crossentropy",
    optimizer = "adam",
    metrics = "accuracy"
  )
```

```{r, cache = TRUE, warning = FALSE}
mnist.model %>%
  fit(
    x = mnist.train$x, y = mnist.train$y,
    epochs = 5,
    validation_split = 0.3,
    verbose = 2
  )
```
```{r, cache = TRUE}
mnist.predictions <- predict(mnist.model, mnist.test$x)
```

```{r, cache = TRUE}
mnist.model %>% 
  evaluate(mnist.test$x, mnist.test$y, verbose = 0)
```
# Fashion MNIST

```{r, cache = TRUE}
fashion <- dataset_fashion_mnist()
```

Taking a look at an example of the images:

```{r, error = FALSE, message = FALSE, warning = FALSE, cache = TRUE}
# Plotting helper; fashion is global var
fashionplot <- function(picked){
  fashion$train$x[picked,,] %>% 
  # Wrangle
  as_tibble() %>% 
  rowid_to_column("Y") %>% 
  pivot_longer(-Y,names_to = "X", values_to = "PIX") %>% 
  mutate(X = as.numeric(gsub("V","",X))) %>%
  mutate(Y = abs(Y - max(Y)+1)) %>%
  # Graph
  ggplot(aes(X,Y, fill = PIX)) + 
  geom_tile() + 
  scale_fill_gradient(low="white", high="black", guide = FALSE) +
  ggtitle(paste("Clothing Category:",fashion$train$y[picked])) +
  theme_void()
}
# Grid graph
pickcount <- 9
fashionex <- lapply(1:pickcount, fashionplot)
fashionexplot <- arrangeGrob(grobs = fashionex,
                           ncol = 3,
                           nrow = 3)
plot(fashionexplot)
```

```{r}
fashion.test = fashion$test
fashion.train = fashion$train
# Normalization; Speeds training time
fashion.train$x <- mnist.train$x/255
fashion.test$x  <- mnist.test$x/255
```

# CIFAR-10

```{r, cache = TRUE}
cifar10 <- dataset_cifar10()
```

Taking a look at CIFAR-10 examples is a little trickier as it is a three-dimensional array.  For this we will just utilize the `imager` library:

```{r}
library(imager)
par(mfrow = c(2,2))
for(i in 1:4){
  plot(cifar10$train$x[i,,,] %>%
         as.cimg %>%
         imrotate(,angle = 90),
       main = cifar10$test$y[i],
       axes = FALSE)
}
```


